---
title: "Animate networks"
author: "Chenxin Li"
date: "2023-01-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
Let's make animated network graphs! 

# Packages 
```{r, echo=FALSE, results='hide'} 
library(tidyverse)
library(igraph)
library(ggraph)

library(gganimate)
library(readxl)
library(RColorBrewer)
```

To do network stuff, we need `igraph` and `ggraph`. 
`igraph` is a network analysis package that can do a lot of heavy-lifting when it comes of dealing with networks. 
`ggraph` is a `ggplot` extension for network visualization.  

To animate ggplot objects, we need gganime. 

# Load an example network 
## Make a network object from edges and nodes 
We will use an example. Here is a hypothetical network that I made up. 
```{r}
my_network_edges <- read_excel("../Data/Example_network_edges.xlsx")

my_network_edges
```
Without a given node table, 
you can actually produce a node table from the edge table. 
The node table is just a non-redundant list of all members of `from` and `to` in the edge table. 

```{r}
my_network_nodes <- data.frame(
  nodes = unique(c(my_network_edges$From, my_network_edges$To))
) %>% 
  #mutate(nodes = as.character(nodes)) %>% 
  mutate(module = case_when(
   nodes %in% c(1:10) ~ "Module 1",
   nodes %in% c(11:20) ~ "Module 2",
   nodes %in% c(21:30) ~ "Module 3",
  ))

my_network_nodes
```

`unique(c(example_network_edges$From, example_network_edges$To)` takes the members of `from` and `to` from `example_network_edges` and take the non-redundant set. 

To make a network, use the `graph_from_data_frame()` function from the `igraph` package. 
```{r}
my_network <- graph_from_data_frame(
  d = my_network_edges,
  vertices = my_network_nodes,
  directed = F
)

my_network
```

The `d` in `graph_from_data_frame()` specifies the edge table.
`vertices` species the node table. 
In this example I set `directed` to `FALSE`.
This means from node 1 to node 2 is the same from node 2 to node 1. 

## Using ggraph 
To make a network diagram, we will use `ggraph`.
```{r}
kk_vis <- ggraph(my_network, 
       layout = "kk"
       ) +
  geom_edge_link(color = "grey70") +
  geom_node_point(size = 3, shape = 21, color = "white", aes(fill = module)) +
  geom_node_text(repel = T, aes(label = my_network_nodes$nodes)) +
  scale_fill_manual(values = brewer.pal(8, "Set2")) +
  theme_void()
```
There are 3 fundamental elements to a network diagram. 

1. Layout, which controls how nodes are placed in the 2-D plane. In this diagram I used "kk" or "Kamada-Kawai" layout algorithm. More about layout algorithms is discussed below. 
2. Edges. In this example, I used `geom_edge_diagonal()`, which draws curves between nodes.
3. Nodes, which is provided by `geom_node_point()`, draws each node as a point. 

# Try different network layouts
There are multiple layout algorithms for network diagrams. 
Layouts can drastically change the appearance of networks, making them easier or harder to interpret. 


My go-to is "kk", which generally gives good results.
But I encourage you to read more about different layouts [here](https://www.data-imaginist.com/2017/ggraph-introduction-layouts/) and [here](https://r-graph-gallery.com/247-network-chart-layouts.html). 

## Pull out x & y coordinates
```{r}
circle_xy <- ggraph(my_network, layout = "circle")$data %>% 
  mutate(layout = "circle")

star_xy <- ggraph(my_network, layout = "star")$data %>% 
  mutate(layout = "star")

kk_xy <- ggraph(my_network, layout = "kk")$data %>% 
  mutate(layout = "kk")

mds_xy <- ggraph(my_network, layout = "mds")$data %>% 
  mutate(layout = "mds")

fr_xy <- ggraph(my_network, layout = "fr")$data %>% 
  mutate(layout = "fr")

gem_xy <- ggraph(my_network, layout = "gem")$data %>% 
  mutate(layout = "gem")

stress_xy <- ggraph(my_network, layout = "stress")$data %>% 
  mutate(layout = "stress")

tree_xy <- ggraph(my_network, layout = "tree")$data %>% 
  mutate(layout = "tree")

sugiyama_xy <- ggraph(my_network, layout = "sugiyama")$data %>% 
  mutate(layout = "sugiyama")

node_xy <- rbind(
  circle_xy,
  star_xy,
  kk_xy,
  mds_xy,
  fr_xy,
  stress_xy,
  gem_xy,
  tree_xy,
  sugiyama_xy
) %>% 
   mutate(module = case_when(
   name %in% c(1:10) ~ "Module 1",
   name %in% c(11:20) ~ "Module 2",
   name %in% c(21:30) ~ "Module 3",
  )) %>% 
  select(x, y, name, layout, module)

head(node_xy)
```


## Generate the edge data frame

```{r}
circle_edges <- my_network_edges %>% 
  mutate(edge_id = paste0(From, "-", To)) %>% 
  mutate(From = as.character(From)) %>% 
  mutate(To = as.character(To)) %>% 
  inner_join(circle_xy %>% 
               select(x, y, name, layout), by = c("From" = "name")) %>% 
  rename(
    temx = x, 
    temy = y,
  ) %>% 
  inner_join(circle_xy %>% 
               select(x, y, name, layout), by = c("To" = "name", "layout")) %>% 
  rename(
    x = temx,
    y = temy,
    xend = x,
    yend = y
  )

star_edges <- my_network_edges %>% 
  mutate(edge_id = paste0(From, "-", To)) %>% 
  mutate(From = as.character(From)) %>% 
  mutate(To = as.character(To)) %>% 
  inner_join(star_xy %>% 
               select(x, y, name, layout), by = c("From" = "name")) %>% 
  rename(
    temx = x, 
    temy = y,
  ) %>% 
  inner_join(star_xy %>% 
               select(x, y, name, layout), by = c("To" = "name", "layout")) %>% 
  rename(
    x = temx,
    y = temy,
    xend = x,
    yend = y
  )

kk_edges <- my_network_edges %>% 
  mutate(edge_id = paste0(From, "-", To)) %>% 
  mutate(From = as.character(From)) %>% 
  mutate(To = as.character(To)) %>%
  inner_join(kk_xy %>% 
               select(x, y, name, layout), by = c("From" = "name")) %>% 
  rename(
    temx = x, 
    temy = y,
  ) %>% 
  inner_join(kk_xy %>% 
               select(x, y, name, layout), by = c("To" = "name", "layout")) %>% 
  rename(
    x = temx,
    y = temy,
    xend = x,
    yend = y
  )

mds_edges <- my_network_edges %>% 
  mutate(edge_id = paste0(From, "-", To)) %>% 
  mutate(From = as.character(From)) %>% 
  mutate(To = as.character(To)) %>%
  inner_join(mds_xy %>% 
               select(x, y, name, layout), by = c("From" = "name")) %>% 
  rename(
    temx = x, 
    temy = y,
  ) %>% 
  inner_join(mds_xy %>% 
               select(x, y, name, layout), by = c("To" = "name", "layout")) %>% 
  rename(
    x = temx,
    y = temy,
    xend = x,
    yend = y
  )

fr_edges <- my_network_edges %>% 
  mutate(edge_id = paste0(From, "-", To)) %>% 
  mutate(From = as.character(From)) %>% 
  mutate(To = as.character(To)) %>%
  inner_join(fr_xy %>% 
               select(x, y, name, layout), by = c("From" = "name")) %>% 
  rename(
    temx = x, 
    temy = y,
  ) %>% 
  inner_join(fr_xy %>% 
               select(x, y, name, layout), by = c("To" = "name", "layout")) %>% 
  rename(
    x = temx,
    y = temy,
    xend = x,
    yend = y
  )

stress_edges <- my_network_edges %>% 
  mutate(edge_id = paste0(From, "-", To)) %>% 
  mutate(From = as.character(From)) %>% 
  mutate(To = as.character(To)) %>% 
  inner_join(stress_xy %>% 
               select(x, y, name, layout), by = c("From" = "name")) %>% 
  rename(
    temx = x, 
    temy = y,
  ) %>% 
  inner_join(stress_xy %>% 
               select(x, y, name, layout), by = c("To" = "name", "layout")) %>% 
  rename(
    x = temx,
    y = temy,
    xend = x,
    yend = y
  )

gem_edges <- my_network_edges %>% 
  mutate(edge_id = paste0(From, "-", To)) %>% 
  mutate(From = as.character(From)) %>% 
  mutate(To = as.character(To)) %>% 
  inner_join(gem_xy %>% 
               select(x, y, name, layout), by = c("From" = "name")) %>% 
  rename(
    temx = x, 
    temy = y,
  ) %>% 
  inner_join(gem_xy %>% 
               select(x, y, name, layout), by = c("To" = "name", "layout")) %>% 
  rename(
    x = temx,
    y = temy,
    xend = x,
    yend = y
  )

tree_edges <- my_network_edges %>% 
  mutate(edge_id = paste0(From, "-", To)) %>% 
  mutate(From = as.character(From)) %>% 
  mutate(To = as.character(To)) %>%
  inner_join(tree_xy %>% 
               select(x, y, name, layout), by = c("From" = "name")) %>% 
  rename(
    temx = x, 
    temy = y,
  ) %>% 
  inner_join(tree_xy %>% 
               select(x, y, name, layout), by = c("To" = "name", "layout")) %>% 
  rename(
    x = temx,
    y = temy,
    xend = x,
    yend = y
  )

sugiyama_edges <- my_network_edges %>% 
  mutate(edge_id = paste0(From, "-", To)) %>% 
  mutate(From = as.character(From)) %>% 
  mutate(To = as.character(To)) %>% 
  inner_join(sugiyama_xy %>% 
               select(x, y, name, layout), by = c("From" = "name")) %>% 
  rename(
    temx = x, 
    temy = y,
  ) %>% 
  inner_join(sugiyama_xy %>% 
               select(x, y, name, layout), by = c("To" = "name", "layout")) %>% 
  rename(
    x = temx,
    y = temy,
    xend = x,
    yend = y
  )

edge_data_frame <- rbind(
  circle_edges,
  star_edges,
  kk_edges,
  mds_edges,
  fr_edges,
  stress_edges,
  gem_edges,
  tree_edges,
  sugiyama_edges
) 

head(edge_data_frame)
```

 

# Make the animation 
Each frame is a layout 
```{r}
ggplot() +
  geom_segment(
    data = edge_data_frame %>% 
      mutate(layout = factor(layout, levels = c(
        "circle", "star", "tree", 
        "sugiyama", "mds", "fr",
        "kk", "gem", "stress"
      ))),
    aes(x = x, xend = xend, y = y, yend = yend, group = edge_id),
    color = "grey70", linewidth = 1, 
  ) +
  geom_point(
    data = node_xy %>% 
      mutate(layout = factor(layout, levels = c(
        "circle", "star", "tree", 
        "sugiyama", "mds", "fr",
        "kk", "gem", "stress"
      ))), aes(x, y, fill = module),
    shape = 21, size = 4, color = "white", alpha = 0.8
  ) +
  labs(title = "layout: {closest_state}",
       fill = NULL) +
  scale_fill_manual(values = brewer.pal(8, "Set2")) +
  ease_aes("linear") +
  transition_states(layout, state_length = 2, transition_length = 1) + 
  theme_void() +
  theme(
    plot.title = element_text(size = 20),
    legend.position = "none"
  ) +
  view_follow()

anim_save("../Results/network_layouts.gif")
```

```{r}
ggplot() +
  facet_wrap(~layout, scales = "free") +
  geom_segment(
    data = edge_data_frame %>% 
      mutate(layout = factor(layout, levels = c(
        "circle", "star", "tree", 
        "sugiyama", "mds", "fr",
        "kk", "gem", "stress"
      ))),
    aes(x = x, xend = xend, y = y, yend = yend, group = edge_id),
    color = "grey70", 
  ) +
  geom_point(
    data = node_xy %>% 
      mutate(layout = factor(layout, levels = c(
        "circle", "star", "tree", 
        "sugiyama", "mds", "fr",
        "kk", "gem", "stress"
      ))), aes(x, y, fill = module),
    shape = 21, size = 3, color = "white", alpha = 0.8
  ) +
  labs(title = "Different layouts of the SAME network\n",
       fill = NULL,
       x = NULL,
       y = NULL) +
  scale_fill_manual(values = brewer.pal(8, "Set2")) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12),
    legend.position = "none",
    panel.spacing = unit(1, "lines"),
    panel.background = element_rect(fill = NA, color = NA),
    strip.background = element_rect(fill = NA, color = NA),
    panel.grid = element_blank(),
    axis.text = element_blank(),
    strip.text = element_text(size = 12)
  )

ggsave("../Results/Different_layouts.svg", height = 6, width = 6, bg = "white")
ggsave("../Results/Different_layouts.png", height = 6, width = 6, bg = "white")
```

