---
title: "BarMeadow"
author: "Chenxin Li"
date: '2022-05-24'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is an Rmarkdown document.If you don't have the rmarkdown package, you will have to install it to run this. 
The grey areas are code chunks. To run a code chunk click the green triangle at the upper right of each chunk. 

# Packages 
```{r}
library(tidyverse) 
library(ggbeeswarm) # to make jitter plots  
library(readxl) # to load excel files 

library(viridis) # for the colors 
library(patchwork)

library(emmeans) # for estimated marginal means and statistics 
```

You will also need `multcomp` and `multcompView` packages for compact letter display (CLD). 
Install them if you don't have them. 

# Data 

Data from:  
    Matand, K., Shoemake, M. and Li, C., 2020. 
    High frequency in vitro regeneration of adventitious shoots in daylilies (Hemerocallis sp) stem tissue using thidiazuron. 
    BMC Plant Biology, 20(1), pp.1-10.
  
For simiplicity, we will be only three of the 19 cultivars: Alias, Bright, and Rococo

```{r}
stem_data <- read_excel("../Data/Matand_2020_Stem_Data.xlsx") 

stem_data2 <- stem_data %>% 
  mutate(Treatment = factor(Treatment, levels = c("T1", "T5", "T10"))) %>% 
  filter(str_detect(Variety, "Alias|Bright|Rococo"))

head(stem_data)
```


# Stat
```{r}
model_stem <- lm(Primordia ~ Explant * Treatment * Variety, data = stem_data2)
anova(model_stem)
```
Run a quick ANOVA to check which factors/interactions account for the most variances. 

# Bad example
## CLD  
For the classic bar plot meadow, let's generate all pairwise comparisons.
This is indicated by the ~ Variety:Explant:Treatment syntax. 

```{r}
est_all <- emmeans(model_stem, pairwise ~ Variety:Explant:Treatment)    
results_all <- multcomp::cld(est_all$emmeans, Letters = "abcdefghijklmnopqrst") %>% 
  as.data.frame() %>% 
  mutate(grouping = str_remove_all(.group, " "))

head(results_all)
```


## Plot 
The classic bar plot meadow ususally only plots mean and se. 
We will generate that first. 

```{r}
barplotMeadow <- stem_data2 %>% 
  group_by(Variety, Treatment, Explant) %>% 
  summarise(
    mean = mean(Primordia),
    sd = sd(Primordia),
    NN = n()
  ) %>% 
  ungroup() %>% 
  mutate(se = sd/sqrt(NN)) %>% 
  ggplot(aes(x = interaction(Variety, Treatment, Explant), y = mean)) +
  geom_bar(stat = "identity", position = position_dodge2(), alpha = 0.8, width = 0.7) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.2) +
  geom_text(data = results_all, aes(label = grouping, y = upper.CL), vjust = -0.2) +
  labs(x = NULL,
       y = "Mean",
       caption = "This is horrendous.\nWhat am I looking at?") +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.caption = element_text(hjust = 0, size = 12)
  ) +
  ggtitle("Classic bar plot meadow\ny ~ Variety:Treatment:Explant")

barplotMeadow
```

# Good example 
## CLD

In this example, we will focus on comparing the effect of explant and treatment combination at the level of each cultivar. 
This is indicated by the ~ Explant:Treatment | Variety syntax. 

You can also compare the effect of treatment and cultivar at the level of each explant, if that's what you are interested in. 
In that case, you will need ~ Treatment:Variety | Explant. 

```{r}
est_by_factor <- emmeans(model_stem, pairwise ~ Explant:Treatment | Variety)
results_by_factor <- multcomp::cld(est_by_factor$emmeans, Letters = "abcdefghijklmnopqrst") %>% 
  as.data.frame() %>% 
  mutate(grouping = str_remove_all(.group, " "))

head(results_by_factor)
```

 
## plot
```{r}
DotPlot_by_factor <- stem_data2 %>% 
  ggplot(aes(x = Treatment, y = Primordia)) +
  facet_wrap( ~ Variety + Explant, scales = "free_x", strip.position = "bottom") +
  ggbeeswarm::geom_quasirandom(aes(fill = Treatment), 
                               shape = 21, color = "white", size = 2.5, alpha = 0.8) +
  geom_text(data = results_by_factor, 
            aes(label = grouping, y = upper.CL + 35), size = 5) +
  scale_fill_manual(values = viridis(10)[c(1, 5, 8)]) +
  labs(y = "Response",
       x = NULL,
       caption = "That's better.\nReader's attention is more focused.") +
  theme_classic() +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    panel.spacing.y = unit(1.2, "lines"),
    strip.text = element_text(size = 10),
    strip.placement = "outside",
    plot.caption = element_text(hjust = 0, size = 12)
  ) +
  ggtitle("Better designed for multifactorial experiments\ny ~ Treatment:Explant | Variety")

DotPlot_by_factor
```
This has variety as rows and explant as columns, which allows us to compare across the same variety. 
Again, this is indicated by ~ Explant:Treatment | Variety syntax. 


# Put together
```{r}
wrap_plots(barplotMeadow, DotPlot_by_factor, nrow = 1)

ggsave("../Results/AvoidBarPlotMeadow.svg", height = 7, width = 10, bg = "white")
ggsave("../Results/AvoidBarPlotMeadow.png", height = 7, width = 10, bg = "white")
```






