---
title: "Bar_charts"
author: "Chenxin Li"
date: "4/5/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is an Rmarkdown document.If you don't have the rmarkdown package, you will have to install it to run this. 
The grey areas are code chunks. To run a code chunk click the green triangle at the upper right of each chunk.  

# Package
```{r}
library(tidyverse)
library(ggbeeswarm) # to make jitter plots  
library(RColorBrewer) # for the colors 
library(patchwork) # for putting ggplot objects into one graph 
```

# Data

Group is sampled from a normal distribution with mean = 1 and sd = 1. 
Group is sampled from a lognormal distribution with mean = 1 and sd = 1. 

```{r}
set.seed(666)
group1 <- rnorm(n = 100, mean = 1, sd = 1)
group2 <- rlnorm(n = 100, 
                 meanlog = log(1^2/sqrt(1^2 + 1^2)), 
                 sdlog = sqrt(log(1+(1^2/1^2))))

sd(group1)
sd(group2)

mean(group1)
mean(group2)

median(group1)
median(group2)

IQR(group1)
IQR(group2)
```

```{r}
groups_long <- cbind(
  group1,
  group2
) %>% 
  as.data.frame() %>% 
  gather("group", "response", 1:2)

head(groups_long)
```

# Tests
```{r}
t.test(group1, group2)

wilcox.test(group1, group2)

ks.test(group1, group2)
```

You will notice different tests give different results. 
Might seem odd at first, but these tests are not testing for the same thing, so it is not odd after all.

# Plot
## bar
```{r}
bar <- groups_long %>% 
  ggplot(aes(x = group, y = response)) +
  geom_bar(stat = "summary", fun = mean, 
           width = 0.7, alpha = 0.8,
           aes(fill = group)) + 
  stat_summary(geom = "errorbar", fun.data = "mean_se",
               width = 0.1, size = 1) +
  scale_fill_manual(values = brewer.pal(8, "Accent")[1:2]) +
  labs(x = "Group",
       y = "Response",
       caption = paste0("They are the same!\nP = ", 
                        signif(t.test(group1, group2)$p.value, 2),
                        " (t test)")) +
  theme_classic() +
  theme(
    text = element_text(size = 12, face = "bold", color = "black"),
    axis.text = element_text(color = "black"),
    legend.position = "none",
    plot.title = element_text(size = 10),
    plot.caption = element_text(hjust = 0)
  ) +
  ggtitle(
    paste0(
      "group1: mean = ", signif(mean(group1), 2), 
      "; sd = ", signif(sd(group1), 2), "\n",
      "group2: mean = ", signif(mean(group2), 2), 
      "; sd = ", signif(sd(group2), 2)
    )) 
  
  

bar
```
## Boxplot
```{r}
box <- groups_long %>% 
  ggplot(aes(x = group, y = response)) +
  geom_boxplot(width = 0.7, alpha = 0.8, 
          aes(fill = group)) +
  scale_fill_manual(values = brewer.pal(8, "Accent")[1:2]) +
  labs(x = "Group",
       y = "Response",
       caption = paste0("Hmmmmm...\nP = ", 
                        signif(wilcox.test(group1, group2)$p.value, 2),
                        " (Wilcoxon rank sum test)")) +
  theme_classic() +
  theme(
    text = element_text(size = 12, face = "bold", color = "black"),
    axis.text = element_text(color = "black"),
    legend.position = "none",
    plot.title = element_text(size = 10),
    plot.caption = element_text(hjust = 0)
  ) +
  ggtitle(
    paste0(
      "group1: median = ", signif(median(group1), 2), 
      "; IQR = ", signif(IQR(group1), 2), "\n",
      "group2: median = ", signif(median(group2), 2), 
      "; IQR = ", signif(IQR(group2), 2)
    )
  )

box
```
## Dot plot
```{r}
dotplot <- groups_long %>% 
  ggplot(aes(x = group, y = response)) +
  ggbeeswarm::geom_quasirandom(
    shape = 21, color = "white", 
    alpha = 0.8, size = 3,
    aes(fill = group)
  ) +
  scale_fill_manual(values = brewer.pal(8, "Accent")[1:2]) +
  labs(x = "Group",
       y = "Response",
        caption = paste0("OH!!!\nP = ", 
                        signif(ks.test(group1, group2)$p.value, 2),
                        " (Kolmogorov–Smirnov test)")) +
  theme_classic() +
  theme(
    text = element_text(size = 12, face = "bold", color = "black"),
    axis.text = element_text(color = "black"),
    legend.position = "none",
    plot.title = element_text(size = 10),
    plot.caption = element_text(hjust = 0)
  ) +
  ggtitle(
    paste0(
      "group1: median = ", signif(median(group1), 2), 
      "; IQR = ", signif(IQR(group1), 2), "\n",
      "group2: median = ", signif(median(group2), 2), 
      "; IQR = ", signif(IQR(group2), 2)
    )
  )

dotplot 
```
# combine plots
```{r}
wrap_plots(
  bar, box, dotplot, nrow = 1
)

ggsave("../Results/dont_bar_plot.svg", height = 5, width = 10)
ggsave("../Results/dont_bar_plot.png", height = 5, width = 10)
```

