---
title: "BoxPlot_for_Binomial"
author: "Chenxin Li"
date: "2024-12-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Friends don't let friends use boxpot for binomial data

# Packages
```{r}
library(tidyverse)
library(RColorBrewer)
library(ggbeeswarm)

library(patchwork)
```

## Data 
```{r}
set.seed(666)
```

```{r}
data1 <- data.frame(
  response = rnorm(n = 100, mean = 5, sd = 2)
) %>% 
  mutate(group = "group1")

data2 <- data.frame(
  response = c(
    rnorm(n = 50, mean = 2.5, sd = 1),
    rnorm(n = 50, mean = 7.5, sd = 1)
  )) %>% 
    mutate(group = "group2")

data3 <- data.frame(
  response = c(
    rnorm(n = 33, mean = 2, sd = 0.5),
    rnorm(n = 33, mean = 5, sd = 0.5),
    rnorm(n = 33, mean = 8, sd = 0.5)
  )) %>% 
    mutate(group = "group3")
```

## Bad example
```{r}
Box <- rbind(
  data1, 
  data2,
  data3
) %>% 
  ggplot(aes(x = group, y = response)) +
  geom_boxplot(aes(fill = group), alpha = 0.8, width = 0.7) +
  scale_fill_manual(values = brewer.pal(8, "Set2")) +
  labs(title = "Very similar!") + 
  theme_classic()

Box
```
## Good example
```{r}
Dots <- rbind(
  data1, 
  data2,
  data3
) %>% 
  ggplot(aes(x = group, y = response)) +
  geom_quasirandom(aes(color = group), alpha = 0.8) +
  scale_color_manual(values = brewer.pal(8, "Set2")) +
  labs(title = "I guess not!") + 
  theme_classic()

Dots
```
# Wrap them together
```{r}
wrap_plots(
  Box, Dots,
  nrow = 1
) &
  theme(legend.position = "none")

ggsave("../Results/BoxPlots_for_binomial.svg", width = 5, height = 2.5)
ggsave("../Results/BoxPlots_for_binomial.png", width = 5, height = 2.5)
```
Before making a box plot, one should check the distribution of their data,
since box plots focus on median and quartiles, 
they cannot handle binomial data (and by extension data with multiple modes).
Ploting all the data points using `geom_quasirandom()` from the [ggbeeswarm package](https://github.com/eclarke/ggbeeswarm) is the best practice for small sample to moderate (less than tens of thousands) sample sizes, 
as dots are robust to small sample sizes, 
whereas distribution-based graphics such as violin plots and histograms are not.
See [this section](https://github.com/cxli233/FriendsDontLetFriends#2-friends-dont-let-friends-make-violin-plots-for-small-sample-sizes) and [this section](https://github.com/cxli233/FriendsDontLetFriends/tree/main?tab=readme-ov-file#friends-dont-let-friends-use-histogram-for-small-sample-sizes) for details. 