---
title: "Colorscales"
author: "Chenxin Li"
date: '2022-05-06'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is an Rmarkdown document.If you don't have the rmarkdown package, you will have to install it to run this. 
The grey areas are code chunks. To run a code chunk click the green triangle at the upper right of each chunk. 

# Packages 
```{r}
library(tidyverse) 
library(ggbeeswarm) # to make jitter plots 
library(RColorBrewer) # for colors 
library(viridis) # for colors 
library(patchwork) # for putting graphs together 
```

# Simulate data 
Just some arbitrary sequential/directional data 
```{r}
Dir1 <- data.frame(
  "FPKM" = c(5, 10, 20, 40, 60, 80, 100, 120),
  "Gene" = 1:8
)  

Dir2 <- data.frame(
  "log2FC" = c(-10, -4, -2, 0, 2, 4, 10),
  "Gene" = 1:7
)

Dir3 <- data.frame(
  "z.score" = c(-2.5, -2, -1, 0, 1, 2, 2.5),
  "Gene" = 1:7
)
  
head(Dir3)
```

# Graph
## Good example 1 
```{r}
A <- Dir1 %>% 
 ggplot(aes(x = Gene, y = FPKM)) +
  geom_hline(yintercept = 0, size = 1) +
  geom_segment(aes(x = Gene, xend = Gene), 
               yend = 0, size = 1.2, alpha = 0.8, color = "grey60") +
  geom_point(shape = 21, size = 4, aes(fill = FPKM)) +
  scale_fill_gradientn(colors = viridis(n = 10)) +
  labs(caption = "This is good.") +
  theme_minimal() +
  theme(
    text = element_text(size = 14, color = "black"), 
    axis.text = element_text(color = "black"), 
    axis.text.y = element_blank(), 
    legend.position = "none",
    plot.caption = element_text(hjust = 0, size = 12)
  ) +
  coord_flip()+
  ggtitle("Darkest color = Min\nLightest color = Max")

B <- Dir1 %>% 
  ggplot(aes(x = " ", y = Gene)) +
  geom_tile(aes(fill = FPKM)) +
  scale_fill_gradientn(colors = viridis(n = 10)) +
  theme_void() +
  theme(
    legend.position = "none"
  )

Good1 <- wrap_plots(A, B, nrow = 1, widths = c(1, 0.1))

Good1
```
This is a good example, where lightest color = max, and darkest color = min. 


## Good example 2 
```{r}
C <- Dir2 %>% 
 ggplot(aes(x = Gene, y = log2FC)) +
  geom_hline(yintercept = 0, size = 1) +
  geom_segment(aes(x = Gene, xend = Gene), 
               yend = 0, size = 1.2, alpha = 0.8, color = "grey60") +
  geom_point(shape = 21, size = 4, aes(fill = log2FC)) +
  scale_fill_gradientn(colors = rev(brewer.pal(11, "RdBu"))) +
  labs(caption = "This is good.") +
  theme_minimal() +
  theme(
    text = element_text(size = 14, color = "black"), 
    axis.text = element_text(color = "black"), 
    axis.text.y = element_blank(), 
    legend.position = "none",
    plot.caption = element_text(hjust = 0, size = 12)
  ) +
  coord_flip()+
  ggtitle("Lightest color = 0\nDarkest colors = Max absolutes")

D <- Dir2 %>% 
  ggplot(aes(x = " ", y = Gene)) +
  geom_tile(aes(fill = log2FC)) +
  scale_fill_gradientn(colors = rev(brewer.pal(11, "RdBu"))) +
  theme_void() +
  theme(
    legend.position = "none"
  )

Good2 <- wrap_plots(C, D, nrow = 1, widths = c(1, 0.1))

Good2
```
This is a good example, where lightest color = 0 and have a special meaning, and darkest colors = max absolute values.  

## Good example 3
```{r}
E <- Dir3 %>% 
 ggplot(aes(x = Gene, y = z.score)) +
  geom_hline(yintercept = 0, size = 1) +
  geom_segment(aes(x = Gene, xend = Gene), 
               yend = 0, size = 1.2, alpha = 0.8, color = "grey60") +
  geom_point(shape = 21, size = 4, aes(fill = z.score)) +
  scale_fill_gradientn(colors = brewer.pal(9, "YlGnBu")) +
  labs(caption = "This is good.",
       y = "z score") +
  theme_minimal() +
  theme(
    text = element_text(size = 14, color = "black"), 
    axis.text = element_text(color = "black"), 
    axis.text.y = element_blank(), 
    legend.position = "none",
    plot.caption = element_text(hjust = 0, size = 12)
  ) +
  coord_flip()+
  ggtitle("Darkest color = Max\nLightest color = Min")

eF <- Dir3 %>% 
  ggplot(aes(x = " ", y = Gene)) +
  geom_tile(aes(fill = z.score)) +
  scale_fill_gradientn(colors = brewer.pal(9, "YlGnBu")) +
  theme_void() +
  theme(
    legend.position = "none"
  )

Good3 <- wrap_plots(E, eF, nrow = 1, widths = c(1, 0.1))

Good3
```

This is a good example, where lightest color = min, and darkest colors = max. 

## Bad example 
```{r}
G <- Dir1 %>% 
 ggplot(aes(x = Gene, y = FPKM)) +
  geom_hline(yintercept = 0, size = 1) +
  geom_segment(aes(x = Gene, xend = Gene), 
               yend = 0, size = 1.2, alpha = 0.8, color = "grey60") +
  geom_point(shape = 21, size = 4, aes(fill = FPKM)) +
  scale_fill_gradientn(colors = rev(brewer.pal(11, "RdBu"))) +
  labs(caption = "A data visualization sin") +
  theme_minimal() +
  theme(
    text = element_text(size = 14, color = "black"), 
    axis.text = element_text(color = "black"), 
    axis.text.y = element_blank(), 
    legend.position = "none",
    plot.caption = element_text(hjust = 0, size = 12)
  ) +
  coord_flip()+
  ggtitle("Lighest color means nothing\n(neither mean nor median).")

H <- Dir1 %>% 
  ggplot(aes(x = " ", y = Gene)) +
  geom_tile(aes(fill = FPKM)) +
  scale_fill_gradientn(colors = rev(brewer.pal(11, "RdBu"))) +
  theme_void() +
  theme(
    legend.position = "none"
  )

Bad <- wrap_plots(G, H, nrow = 1, widths = c(1, 0.1))

Bad
```

This is a bad example and a data visualization sin. 
Somehow the lightest color does not mean anything special. 

# Put them all together
```{r}
wrap_plots(
  Good1, Good2, Good3, Bad,
  nrow = 2, ncol = 2
)

ggsave("../Results/ColorScales.svg", height = 7.5, width = 7.5, bg = "white")
ggsave("../Results/ColorScales.png", height = 7.5, width = 7.5, bg = "white")
```


