---
title: "Histogram_for_small_n"
author: "Chenxin Li"
date: "2024-12-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Friends don't let friends use histogram for small sample sizes 

# Packages
```{r}
library(tidyverse)
library(RColorBrewer)
library(ggbeeswarm)
library(viridis)
library(patchwork)
```

# Data
```{r}
set.seed(666)
```

```{r}
n10 <- data.frame(
  response = rnorm(n = 10)
) %>% 
  mutate(group = "n = 10")

n100 <- data.frame(
  response = rnorm(n = 100)
) %>% 
  mutate(group = "n = 100")

n1000 <- data.frame(
  response = rnorm(n = 1000)
) %>% 
  mutate(group = "n = 1000")
```

# Graphs
```{r}
bins10 <- rbind(
  n10, n100, n1000
) %>% 
  ggplot(aes(x = response)) +
  facet_wrap(~group, scales = "free", ncol = 1) +
  geom_histogram(bins = 10, width = 0.7, color = "white", alpha = 0.8, 
                 fill = viridis(n = 8, begin = 0.1, end = 0.8)[1]) +
  labs(title = "10 bins") +
  theme_classic() + 
  theme(panel.spacing = unit(1, "lines"),
        strip.placement = "outside",
        strip.background = element_blank(),
        strip.text = element_text(hjust = 0)) 

bins10
```

```{r}
bins30 <- rbind(
  n10, n100, n1000
) %>% 
  ggplot(aes(x = response)) +
  facet_wrap(~group, scales = "free", ncol = 1) +
  geom_histogram(bins = 30, width = 0.7, color = "white",
                 fill = viridis(n = 8, begin = 0.1, end = 0.8)[4]) +
  labs(title = "30 bins") +
  theme_classic() + 
  theme(panel.spacing = unit(1, "lines"),
        strip.placement = "outside",
        strip.background = element_blank(),
        strip.text = element_text(hjust = 0)) 

bins30
```

```{r}
bins50 <- rbind(
  n10, n100, n1000
) %>% 
  ggplot(aes(x = response)) +
  facet_wrap(~group, scales = "free", ncol = 1) +
  geom_histogram(bins = 50, width = 0.7, color = "white",
                 fill = viridis(n = 8, begin = 0.1, end = 0.8)[7]) +
  labs(title = "50 bins") +
  theme_classic() + 
  theme(panel.spacing = unit(1, "lines"),
        strip.placement = "outside",
        strip.background = element_blank(),
        strip.text = element_text(hjust = 0))

bins50
```

# wrap them
```{r}
wrap_plots(
  bins10, bins30, bins50 +
    labs(caption = "\nWow, the appearance does change with different bin numbers."),
  ncol = 3
) &
  theme(plot.caption = element_text(size = 10))

ggsave("../Results/Histogram_for_small_n.svg", height = 6, width = 8)
ggsave("../Results/Histogram_for_small_n.png", height = 6, width = 8)
```
I've seen histogram being proposed as the replacement for bar plots. 
However, a serious caveat for histogram is that histograms are not robust to bin numbers for small (and even moderate) sample sizes. 
What is a histogram anyway? In a histogram, we first bin the data into a defined number of bins.
Then we count how many observations are there for each bin and graph them. 

In this example, I sampled _the same_ normal distribution 3 times with different sample sizes (n = 10, 100, and 1000).
Even though they came from _the same_ normal distribution, the histograms look quite different based on the number of bins. 
To showcase this, I plotted histogram for 10, 30, and 50 bins. 

First of all, histogram makes no sense for small sample sizes. With small sample sizes (n < 30), the better practice is to graph all data points. 
Second of all, you can see that the shape of the histogram is only robust to changing bin number when the sample size is fairly large (like 1000).
Even if n = 100, the appearance of the histogram can change drastically as the number of bins changes. 

