---
title: "Stacked_bar_and_mean_sep"
author: "Chenxin Li"
date: "2024-01-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages
```{r}
library(tidyverse) 

library(emmeans) # for linear model 
library(RColorBrewer) # for the colors 
library(rcartocolor) # for the colors 
library(patchwork) # for putting graphs together
```

# Data 
```{r}
# Simulate some data 
set.seed(666)

control_1 <- rnorm(n = 5, mean = 30, sd = 5) 
control_2 <- rnorm(n = 5, mean = 50, sd = 5)
control_3 <- rnorm(n = 5, mean = 20, sd = 5) 

total_mean <- 100
adjustment_ctrl <- total_mean / sum(c(mean(control_1), mean(control_2), mean(control_3)))
control_1 <- control_1 * adjustment_ctrl
control_2 <- control_2 * adjustment_ctrl
control_3 <- control_3 * adjustment_ctrl

trt_1 <- rnorm(n = 5, mean = 10, sd = 5)
trt_2 <- rnorm(n = 5, mean = 30, sd = 5)
trt_3 <- rnorm(n = 5, mean = 60, sd = 5)

adjustment_trt <- total_mean / sum(c(mean(trt_1), mean(trt_2), mean(trt_3)))
trt_1 <- trt_1 * adjustment_trt
trt_2 <- trt_2 * adjustment_trt
trt_3 <- trt_3 * adjustment_trt
```

Let's say in this hypothetical experiment, we had blueberry plants in two groups.
One group was the control; the other was treated with a chemical to make fruit development faster.  
Each group had 5 plants. 
The response of the treatment was divided into 3 categories: 
1) light green fruits, 2) light blue fruits, and 3) dark blue fruits. 
100 fruits from each plant were examined and the number of fruits in each category was counted. 
The percentage of fruits in each category was calculated and reported. 
The question of the study is: did the chemical treatment work? 

## Data wrangling 
```{r}
my_data <- rbind(
  control_1 %>% 
  as.data.frame() %>% 
  mutate(group = "ctrl") %>% 
  mutate(category = "light green"), 
  
  control_2 %>% 
  as.data.frame() %>% 
  mutate(group = "ctrl") %>% 
  mutate(category = "light blue"),
  
  control_3 %>% 
  as.data.frame() %>% 
  mutate(group = "ctrl") %>% 
  mutate(category = "dark blue"),
  
  trt_1 %>% 
  as.data.frame() %>% 
  mutate(group = "trt") %>% 
  mutate(category = "light green"), 
  
  trt_2 %>% 
  as.data.frame() %>% 
  mutate(group = "trt") %>% 
  mutate(category = "light blue"),
  
  trt_3 %>% 
  as.data.frame() %>% 
  mutate(group = "trt") %>% 
  mutate(category = "dark blue")
  ) %>% 
  rename(percent_of_fruit = ".")

head(my_data)
```

# Stats
## Summary stats 
```{r}
my_data_s <- my_data %>% 
  group_by(group, category) %>% 
  summarise(mean_percentage = mean(percent_of_fruit),
            sd_percentage = sd(percent_of_fruit),
            N = n()) %>% 
  ungroup()

head(my_data_s)
```
## Linear models 
```{r}
my_model1 <- lm(percent_of_fruit ~ group * category, data = my_data)
anova(my_model1)
```
```{r}
my_estimates <- emmeans(my_model1, pairwise ~ group | category)
my_contrasts <- my_estimates$contrasts %>% 
  as.data.frame() %>% 
  mutate(txt_y = case_when(
                category == "light green" ~ 45,
                category == "light blue" ~ 55,
                category == "dark blue" ~ 35
              ))

my_contrasts
```





# Graphs 
## Stacked bar for 
```{r}
stacked_bar_plot <-my_data_s %>% 
  ggplot(aes(x = group, y = mean_percentage)) +
  geom_bar(stat = "identity", aes(fill = category), 
           width = 0.7, color = "white", alpha = 0.8) +
  scale_y_continuous(breaks = c(0, 50, 100)) +
  scale_fill_manual(values = c(
    carto_pal(7, "BluYl")[2], 
    carto_pal(7, "Purp")[2],
    carto_pal(7, "Sunset")[c(7)]
  ), limits = c("light green", "light blue", "dark blue")) +
  labs(x = "group",
       y = "% of fruits",
       fill = "fruit color", 
       caption = "This is fine.") +
  theme_classic() +
  theme(
    plot.caption = element_text(hjust = 0)
  )

stacked_bar_plot
```


## Graph, the wrong way 
```{r}
stacked_bar_plot_bad <- my_data %>% 
  mutate(off_set = case_when(
    group == "ctrl" & 
      category == "light green" ~ 0,
    group == "ctrl" & 
      category == "light blue" ~ as.numeric(my_data_s[3, 3]),
    group == "ctrl" & 
      category == "dark blue" ~ as.numeric(my_data_s[3, 3] + my_data_s[2, 3]),
    group == "trt" & 
      category == "light green" ~ 0,
    group == "trt" & 
      category == "light blue" ~ as.numeric(my_data_s[6, 3]),
    group == "trt" & 
      category == "dark blue" ~ as.numeric(my_data_s[6, 3] + my_data_s[5, 3])
  )) %>% 
  mutate(percent_of_fruit_offset = percent_of_fruit + off_set) %>% 
  ggplot(aes(x = group, y = percent_of_fruit_offset)) +
  geom_bar(data = my_data %>% 
  group_by(group, category) %>% 
  summarise(mean_percentage = mean(percent_of_fruit)) %>% 
  ungroup(),
  stat = "identity", aes(fill = category, y = mean_percentage),
           width = 0.7, color = "white", alpha = 0.8) +
  geom_point(fill = "grey90", size = 2, shape = 21, color = "grey20",
             position = position_jitter(width = 0.1, seed = 666)) +
  stat_summary(fun.data = mean_se, geom = "linerange", 
               aes(group = category), size = 1) +
  scale_y_continuous(breaks = c(0, 50, 100)) +
  scale_fill_manual(values = c(
    carto_pal(7, "BluYl")[2],
    carto_pal(7, "Purp")[2],
    carto_pal(7, "Sunset")[c(7)]
  ), limits = c("light green", "light blue", "dark blue")) +
  labs(x = "group",
       y = "% of fruits",
       fill = "fruit color",
       caption = "What am I comparing?\nSome dots are above 100%?") +
  theme_classic() +
  theme(
    plot.caption = element_text(hjust = 0)
  )

stacked_bar_plot_bad
```

## Mean separation plot, the better way 
```{r}
jitter_plot <- my_data %>% 
  mutate(category = factor(category, levels = c(
    "light green", "light blue", "dark blue"
  ))) %>% 
  ggplot(aes(x = group, y = percent_of_fruit)) +
  facet_grid(.~ category, scales = "free_y") +
  geom_point(aes(fill = category), color = "grey30", size = 2, alpha = 0.8,
             shape = 21, position = position_jitter(width = 0.1, seed = 666)) +
  stat_summary(fun.data = mean_se, geom = "linerange", 
               aes(group = category), size = 1) +
  geom_segment(data = my_contrasts, aes(y = txt_y, yend = txt_y),
               x = 1.2, xend = 1.8) + 
  geom_text(data = my_contrasts, vjust = -0.25, x = 1.5,
            size = 3, 
            aes(label = paste0(
              "p = ", signif(p.value, 2)
            ),
            y = txt_y)) +
  scale_fill_manual(values = c(
    carto_pal(7, "BluYl")[2],
    carto_pal(7, "Purp")[2],
    carto_pal(7, "Sunset")[c(7)]
  ), limits = c("light green", "light blue", "dark blue")) +
  labs(x = "group",
       y = "% of fruits",
       fill = "fruit color",
       caption = "The better way for mean separation.") +
  guides(fill = "none") +
  theme_classic() +
  theme(
    plot.caption = element_text(hjust = 0),
    panel.spacing = unit(0.6, "lines"),
    strip.text = element_blank(),
    strip.background = element_blank()
  )

jitter_plot
```
## wrap plots together 
```{r}
wrap_plots(
  stacked_bar_plot, stacked_bar_plot_bad, jitter_plot, 
  ncol = 3, guides = "collect", 
  widths = c(0.6, 0.6, 1)
) & 
  theme(legend.position = "top")

ggsave("../Results/stacked_bar_vs_jitter.svg", height = 4, width = 8, bg = "white")
ggsave("../Results/stacked_bar_vs_jitter.png", height = 4, width = 8, bg = "white")
```
The first stacked bar plot is fine as the standard way to visualize proportion data. 
It is clear that all categories add up to 100%, 
and the chemical treatment strongly shifted the color profile towards the most developed stage (dark blue). 

The middle stacked bar plot is problematic, 
mainly because it is trying to do two distinct data visualization tasks at once. 
When error bars and dots are overlaid onto the stacked bars, 
it become unclear which error bars and dots are being compared. 
Due to the nature of stacked bars, the error bars and dots of the upper stacks have to be shifted upwards,
and thus interpretation of the y-axis for error bars and dots become not straightforward. 

Finally, if the main point of the visualization is mean separation and dispersion around the mean, 
the third graph is the better choice. 
There is no ambiguity on which comparisons are being made.
As shown in the first stacked bar plot, 
the chemical treatment strongly increases the proportion of dark blue fruits, 
at the expense of lighter color fruits. 


