---
title: "NoConcentricDonunts"
author: "Chenxin Li"
date: '2022-07-28'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages
```{r}
library(tidyverse)
library(RColorBrewer)
library(patchwork) 
```

# Data 
This section generate the data. 
```{r}
group1 <- data.frame(
  "Type" = c("Type I", "Type II"),
  "Percentage" = c(20, 80)
)

group2 <- data.frame(
  "Type" = c("Type I", "Type II"),
  "Percentage" = c(30, 70)
)

group3 <- data.frame(
  "Type" = c("Type I", "Type II"),
  "Percentage" = c(30, 70)
)

blank <- data.frame(
  "Type" = c("Type I", "Type II"),
  "Percentage" = c(0, 0)
)

head(group2)
```

In order for the concentric donuts to work, a "blank" data table is required as the first layer. You will see why. 
 
 
The following section produces `ymax` and `ymin` values for the `geon_rect()` function.
These are required to the concentric donuts. 

```{r}
group1$ymax <- cumsum(group1$Percentage)
group1$ymin <- c(0, head(group1$ymax, n = (nrow(group1)-1)))

group2$ymax <- cumsum(group2$Percentage)
group2$ymin <- c(0, head(group2$ymax, n = (nrow(group2)-1)))

group3$ymax <- cumsum(group3$Percentage)
group3$ymin <- c(0, head(group3$ymax, n = (nrow(group3)-1)))

blank$ymax <- cumsum(blank$Percentage)
blank$ymin <- c(0, head(blank$ymax, n = (nrow(blank)-1)))

head(blank)
```


```{r}
example_data <- rbind(
  blank %>% 
    mutate(group = 0),
  group1 %>% 
    mutate(group = 1),
  group2 %>% 
    mutate(group = 2),
  group3 %>% 
    mutate(group = 3)
)
```


# Graphics
## Bad example 1
```{r}
bad_example_1 <- example_data %>% 
  ggplot(aes(x = group, y = Percentage)) +
  geom_rect(aes(ymax = ymax, ymin = ymin,
                xmax = group-0.25, xmin = group+0.25, 
                fill = Type), alpha = 0.8) +
  annotate(geom = "text", label = c("Group1", "Group2", "Group3"), 
           x = c(1, 2, 3), y = c(0, 0, 0), size = 4) +
  scale_fill_manual(values = brewer.pal(8, "Accent")[c(1, 5)])+
  labs(x = NULL,
       y = NULL,
       fill = NULL,
       title = "Arc lenghts are misleading",
       caption = "Outer rings = much longer arcs") +
  theme_minimal() +
  theme(
    text = element_text(size = 14, color = "black"),
    axis.text = element_text(color = "black"),
    panel.grid = element_blank(),
    axis.text.y = element_blank(),
    legend.position = "top",
    plot.title = element_text(size = 12),
    plot.caption = element_text(hjust = 0)
  ) +
  coord_polar(theta = "y")

bad_example_1
```
You might be tempted to say the data is represented by the arc length, which is in fact inaccurate. 
The arc length on the outer ring is much longer than that in the inner rings. 
Group 2 and Group 3 have the same exact values, but the arc lengths of Group 3 are much longer. 
In fact the data is represented by the arc angle, which we are bad are reading. 
 
## Bad example 2
```{r}
bad_example_2 <- example_data %>% 
  mutate(x = case_when(
    group == 0 ~ 0,
    group == 1 ~ 3,
    group == 2 ~ 2,
    group == 3 ~ 1
  )) %>% 
  mutate(xmax = x - 0.25) %>% 
  mutate(xmin = x + 0.25) %>% 
  ggplot(aes(x = x, y = Percentage)) +
  geom_rect(aes(ymax = ymax, ymin = ymin,
                xmax = xmax, xmin = xmin, 
                fill = Type), alpha = 0.8) +
  annotate(geom = "text", label = c("Group1", "Group2", "Group3"), 
           x = c(3, 2, 1), y = c(0, 0, 0), size = 4) +
  scale_fill_manual(values = brewer.pal(8, "Accent")[c(1,5)])+
  labs(x = NULL,
       y = NULL,
       fill = NULL,
        title = "Ring order matters",
       caption = "Shorter arcs can have larger values") +
  theme_minimal() +
  theme(
    text = element_text(size = 14, color = "black"),
    axis.text = element_text(color = "black"),
    panel.grid = element_blank(),
    axis.text.y = element_blank(),
    legend.position = "top",
    plot.title = element_text(size = 12),
    plot.caption = element_text(hjust = 0)
  ) +
  coord_polar(theta = "y")

bad_example_2
```
As you can see, the ordering of the categories (which group goes to which ring) has a big impact on the look of the graph. 

## Good example 
```{r}
good_example <- example_data %>%
  filter(group != 0) %>% 
  mutate(Type = factor(Type, levels = c("Type II", "Type I"))) %>% 
  ggplot(aes(x = as.factor(group), y = Percentage)) +
  geom_bar(stat = "identity", aes(fill = Type), 
           width = 0.5, alpha = 0.8, color = "white") +
  scale_fill_manual(values = brewer.pal(8, "Accent")[c(1,5)],
                    limits = c("Type I", "Type II"))+
  labs(x = "Group",
       y = "%",
       fill = NULL,
       title = "Just unwrap the donuts!",
       caption = "Simple but better") +
  theme_classic() +
  theme(
    text = element_text(size = 14, color = "black"),
    axis.text = element_text(color = "black"),
    legend.position = "top",
    plot.title = element_text(size = 12),
    plot.caption = element_text(hjust = 0)
  )  

good_example
```

# Put them together 
```{r}
wrap_plots(bad_example_1, bad_example_2, good_example, 
           nrow = 1, widths = c(1, 1, 0.75))

ggsave("../Results/dont_concentric_donuts.svg", height = 4, width = 8, bg = "white")
ggsave("../Results/dont_concentric_donuts.png", height = 4, width = 8, bg = "white")
```

 

