---
title: "RedGreen"
author: "Chenxin Li"
date: "2023-01-16"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Friends don't let friends use red/green or rainbow for heatmaps! 
Deuteranomaly is the most common type of red/green colorblindness, occurring in 1/16 male and 1/256 female. 
Any color scales that uses shades of red and shades of green in the same time would be a problem for a person with red/green colorblindness (third column of figure). 
In addition, red/green and rainbow do not preserve information well at all when printed on black/white (grey scale, second column in figure). 
Many scientific software still use red/green or rainbow as the default color scales, which drives me crazy. 
More "modern" color scales, such as [viridis](https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html) are both colorblind-friendly and grey scale-safe (third row of figure). 
And they look nice too. 



# Package
```{r}
library(tidyverse)
library(viridis)
library(prismatic)
library(patchwork)
```

# Data
```{r}
abc_1 <- expand.grid(
  a = c(1:10),
  b = c(1:10)
) %>% 
  mutate(c = a + b)

head(abc_1)
```
## Write a function to make grey scale 
```{r}
to_grey_scale <- function(hex){
  temp = col2rgb(hex) %>% 
    t() %>% 
    as.data.frame() %>% 
    mutate(gs = 0.299*red + 0.587*green + 0.114*blue)
  
  rgb(temp$gs, temp$gs, temp$gs, maxColorValue = 256)
}
```


# Red/green
```{r}
Red_green <- abc_1 %>% 
  ggplot(aes(x = a, y = b)) +
  geom_tile(aes(fill = c)) +
  scale_fill_gradient2(low = "red", mid = "white", high = "green", midpoint = 11) +
  labs(title = "c = a + b\nred/green\nnormal") +
  theme_classic() +
  theme(legend.key.width = unit(0.7, "lines"),
        axis.text = element_text(color = "black"),
        ) +
  coord_fixed()

Red_green
```
## Grey scale
```{r}
Red_green_gs <- abc_1 %>% 
  ggplot(aes(x = a, y = b)) +
  geom_tile(aes(fill = c)) +
  scale_fill_gradient2(low = to_grey_scale("red"), 
                       mid = "white", 
                       high = to_grey_scale("green"), midpoint = 11) +
  labs(title = "c = a + b\nred/green\ngrey scale") +
  theme_classic() +
  theme(legend.key.width = unit(0.7, "lines"),
        axis.text = element_text(color = "black"),
        ) +
  coord_fixed()

Red_green_gs
```

## Red/green colorblind
```{r}
Red_green_deu <- abc_1 %>% 
  ggplot(aes(x = a, y = b)) +
  geom_tile(aes(fill = c)) +
  scale_fill_gradient2(low = clr_deutan("red"), 
                       mid = clr_deutan("white"), 
                       high = clr_deutan("green"), midpoint = 11) +
  labs(title = "c = a + b\nred/green\ndeuteranopia") +
  theme_classic() +
  theme(legend.key.width = unit(0.7, "lines"),
        axis.text = element_text(color = "black"),
        ) +
  coord_fixed()

Red_green_deu
```


# Rainbow
```{r}
Rainbow1 <- abc_1 %>% 
  ggplot(aes(x = a, y = b)) +
  geom_tile(aes(fill = c)) +
  scale_fill_gradientn(colors = rainbow(20)) +
  labs(title = "c = a + b\nrainbow\nnormal") +
  theme_classic() +
  theme(legend.key.width = unit(0.7, "lines"),
        axis.text = element_text(color = "black"),
        ) +
  coord_fixed()

Rainbow1
```
## grey scale
```{r}
Rainbow_gs <- abc_1 %>% 
  ggplot(aes(x = a, y = b)) +
  geom_tile(aes(fill = c)) +
  scale_fill_gradientn(colors = rainbow(20) %>% to_grey_scale()) +
  labs(title = "c = a + b\nrainbow\ngrey scale") +
  theme_classic() +
  theme(legend.key.width = unit(0.7, "lines"),
        axis.text = element_text(color = "black"),
        ) +
  coord_fixed()

Rainbow_gs
```
## Red/green color blind
```{r}
Rainbow_deu <- abc_1 %>% 
  ggplot(aes(x = a, y = b)) +
  geom_tile(aes(fill = c)) +
  scale_fill_gradientn(colors = rainbow(20) %>% clr_deutan()) +
  labs(title = "c = a + b\nrainbow\ndeuteranopia") +
  theme_classic() +
  theme(legend.key.width = unit(0.7, "lines"),
        axis.text = element_text(color = "black"),
        ) +
  coord_fixed()

Rainbow_deu
```


# Viridis 
## Normal 
```{r}
viridis_optionD <- abc_1 %>% 
  ggplot(aes(x = a, y = b)) +
  geom_tile(aes(fill = c)) +
  scale_fill_gradientn(colors = viridis(100)) +
  labs(title = "c = a + b\nviridis\nnormal") +
  theme_classic() +
  theme(legend.key.width = unit(0.7, "lines"),
        axis.text = element_text(color = "black")) +
  coord_fixed()

viridis_optionD
```
## Grey scale
```{r}
viridis_gs <- abc_1 %>% 
  ggplot(aes(x = a, y = b)) +
  geom_tile(aes(fill = c)) +
  scale_fill_gradientn(colors = to_grey_scale(viridis(100))) +
  labs(title = "c = a + b\nviridis\ngrey scale") +
  theme_classic() +
  theme(legend.key.width = unit(0.7, "lines"),
        axis.text = element_text(color = "black")) +
  coord_fixed()

viridis_gs
```
## Red/green color blind
```{r}
viridis_deu <- abc_1 %>% 
  ggplot(aes(x = a, y = b)) +
  geom_tile(aes(fill = c)) +
  scale_fill_gradientn(colors = clr_deutan(viridis(100))) +
  labs(title = "c = a + b\nviridis\ndeuteranopia") +
  theme_classic() +
  theme(legend.key.width = unit(0.7, "lines"),
        axis.text = element_text(color = "black"),
        ) +
  coord_fixed()

viridis_deu
```
# Put in together: 
```{r}
wrap_plots(
  Red_green, Red_green_gs, Red_green_deu,
  Rainbow1, Rainbow_gs, Rainbow_deu,
  viridis_optionD, viridis_gs, viridis_deu,
  nrow = 3, 
  ncol = 3
)

ggsave("../Results/Color_blind_grey_scale_safe_heatmap.svg", width = 10, height = 7, bg = "white")
ggsave("../Results/Color_blind_grey_scale_safe_heatmap.png", width = 10, height = 7, bg = "white")
```

