---
title: "Reorder_rows_cols_for_heatmap"
author: "Chenxin Li"
date: '2022-06-06'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Friends don't let friends make heatmap without (considering) reordering rows and columns 

# packages 
```{r}
library(tidyverse) 

library(RColorBrewer) # for the colors 
library(viridis) # for the colors 
library(patchwork) # for putting graphs together 
```

# Load data
```{r}
heatmap_data <- read_csv("../Data/deidentified_cell_feature_data.csv")

head(heatmap_data)
```

This is from unpublished data, so I had to de-identify them. 
This is already a tidy data frame, where each row is a cell.

# Check summary 
```{r}
heatmap_data %>% 
  group_by(cell_type) %>% 
  count()
```
The cell types have already been identified - 3 cell types

```{r}
heatmap_data %>% 
  group_by(feature) %>% 
  count() %>% 
  nrow()
```

There are 14 features that were recorded for each cell. 


# Bad example
```{r}
without_reordering <- heatmap_data %>% 
  mutate(z2 = case_when(
    z >= 3 ~ 3,
    T ~ z
  )) %>% 
  ggplot(aes(x = cell, y = feature)) +
  geom_tile(aes(fill = z2)) +
  scale_fill_gradientn(colors = brewer.pal(9, "PuBuGn"), 
                       breaks = c(0, 3),
                       labels = c("0", ">3")) +
  scale_x_discrete(labels = NULL) +
  labs(x = "Cell",
       y = "Feature",
       fill = "z score",
       caption = "Am I looking at a glitching TV?\n",
       title = "Before reordering rows and columns") +
  theme_classic() +
  theme(
    text = element_text(size = 14, color = "black"),
    axis.text = element_text(color = "black"),
    legend.position = "bottom",
    axis.ticks.x = element_blank(),
    plot.caption = element_text(hjust = 0),
    plot.title = element_text(size = 12, face = "bold")
  )

without_reordering
```

# Good example
## Reorder rows and columns 
```{r}
after_reordering <- heatmap_data %>% 
  mutate(z2 = case_when(
    z >= 3 ~ 3,
    T ~ z
  )) %>% 
  mutate(tag = case_when(
    str_detect(feature, " 7") ~ 1,
    str_detect(feature, " 8| 6| 1$") ~ 2,
    str_detect(feature, " 2| 3| 4| 10") ~ 4,
    T ~ 3
  )) %>% 
  mutate(feature = reorder(feature, -tag)) %>% 
  ggplot(aes(x = cell, y = feature)) +
  facet_grid(. ~ cell_type, scales = "free_x", space = "free_x") +
  geom_tile(aes(fill = z2)) +
  scale_fill_gradientn(colors = brewer.pal(9, "PuBuGn"), 
                       breaks = c(0, 3),
                       labels = c("0", ">3")) +
  scale_x_discrete(labels = NULL) +
  labs(x = "Cell",
       y = "Feature",
       fill = "z score",
       caption = "Much better.",
       title = "After reordering rows and columns") +
  theme_classic() +
  theme(
    text = element_text(size = 14, color = "black"),
    axis.text = element_text(color = "black"),
    legend.position = "bottom",
    axis.ticks.x = element_blank(),
    plot.caption = element_text(hjust = 0),
    plot.title = element_text(size = 12, face = "bold")
  )

after_reordering
```

```{r}
wrap_plots(without_reordering, 
           after_reordering, 
           nrow = 2)

ggsave("../Results/Reorder_rows_and_columns_for_heatmap.svg", height = 8, width = 9.5, bg = "white")
ggsave("../Results/Reorder_rows_and_columns_for_heatmap.png", height = 8, width = 9.5, bg = "white")
```




