---
title: "small_n"
author: "Chenxin Li"
date: "4/12/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is an Rmarkdown document.If you don't have the rmarkdown package, you will have to install it to run this. 
The grey areas are code chunks. To run a code chunk click the green triangle at the upper right of each chunk. 

# Packages 
```{r}
library(tidyverse) 
library(ggbeeswarm) # to make jitter plots 

library(emmeans) # for estimated marginal means and stats 

library(RColorBrewer) # for colors 

library(patchwork) # for putting ggplot objects together 
```

# Data
Generate three sets of data from a normal distribution with mean = 1 and sd = 1.

```{r}
set.seed(666) # 

data_set <- replicate(rnorm(5, mean = 1, sd  = 1), n = 3, simplify = T) %>% 
  as.data.frame() %>% 
  gather("sample", "Response", 1:3) %>% 
  mutate(Group = str_replace(sample, "V", "group"))


head(data_set) 
```
# Stats
```{r}
model_1 <- lm(Response ~ Group, data = data_set)
anova(model_1)

emmeans(model_1, ~Group)
```
Group means are clearly not different, as they are sampled from the same distribution. 

# Plot
## Violin 
```{r}
violin_eg <- data_set %>% 
  ggplot(aes(x = Group, y = Response)) +
  geom_violin(aes(fill = Group), 
              alpha = 0.8) +
  stat_summary(geom = "point", fun = median) +
  scale_fill_brewer(palette = "Accent") +
  labs(
    caption = "The distributions are different!\nI wonder what's going on."
  ) +
  theme_classic()+
  theme(
    text = element_text(size=12, color = "black", face = "bold"),
    axis.text = element_text(color = "black"),
    legend.position = "none",
    plot.caption = element_text(hjust = 0)
  ) +
  ggtitle("Points are median.")

violin_eg
```

## Box plot
```{r}
box_eg <- data_set %>% 
  ggplot(aes(x = Group, y = Response)) +
  geom_boxplot(aes(fill = Group), 
              alpha = 0.8) +
  scale_fill_brewer(palette = "Accent") +
  labs(
    caption = "The quartiles are different!\nI wonder what's going on."
  ) +
  theme_classic()+
  theme(
    text = element_text(size=12, color = "black", face = "bold"),
    axis.text = element_text(color = "black"),
    legend.position = "none",
    plot.caption = element_text(hjust = 0)
  ) +
  ggtitle("Boxes span IQR.")

box_eg
```

## jitter plot 
```{r}
jitter_eg <- data_set %>% 
  ggplot(aes(x = Group, y = Response)) +
  geom_point(aes(fill = Group), 
             alpha = 0.8, shape = 21, size = 3, color = "grey40",
             position = position_jitter(seed = 666, width = 0.2)) +
  scale_fill_brewer(palette = "Accent") +
  labs(
    caption = "Never mind...\nToo little data to say anything."
  ) +
  theme_classic()+
  theme(
    text = element_text(size=12, color = "black", face = "bold"),
    axis.text = element_text(color = "black"),
    legend.position = "none",
    plot.caption = element_text(hjust = 0)
  ) +
  ggtitle(
    "n = 5"
  )

jitter_eg
```
### together 
```{r}
wrap_plots(violin_eg, box_eg, jitter_eg,
          nrow = 1)

ggsave("../Results/Beware_of_small_n_box_violin_plot.svg", height = 5, width = 10)
ggsave("../Results/Beware_of_small_n_box_violin_plot.png", height = 5, width = 10)
```

